\textls[-0]{\thetitle}

### 2023-03-04 introduction.tex
intro

The EP is well studied in the conventional setting of functional
programming and object-oriented (OO) programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP.

In contrast, proof assistants offer few linguistic solutions that
address the EP.
Yet, the challenge of writing extensible, type-safe code is
as real, especially for metatheory mechanization.
The programmer faces a tension between adding new constructors to an inductive data type
(e.g., new abstract syntax) and adding new functions and theorems over
the data type (e.g., new meta\-theoretical results).

In the \citeauthor{coq} proof assistant, for instance, inductive types, as well as functions and
theorems over inductive types, are closed to extension.
So to reuse mechanized metatheories,
the common practice is still to copy code and proofs and then modify them in each extension.
But having to maintain multiple copies is highly non-modular and
antithetical to good software engineering.
%
The programmer could also turn to design patterns and encodings~\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
But they tend to require heavy lifting from the programmer to make code
extensible and often lead to non-idiomatic programming styles.

>>>>

The EP is well studied in the setting of ordinary functional
programming and object-oriented (OO) programming.
Modern languages, such as Scala \cite{scala-oopsla05}, have a good
supply of linguistic features that offer expressive power to solve the
EP. %\rD{There is no shortage of solutions to the EP in various
%languages, but there are no citations to those works.}

In contrast, proof assistants offer few linguistic solutions to the EP.
But the challenge of writing extensible, type-safe code is
as real as in any other language.
Metatheory mechanization epitomizes the difficulty:
the programmer faces a tension between adding new constructors to an inductive type
(e.g., new abstract syntax) and adding new functions and theorems over
the inductive type (e.g., new meta\-theoretical results).

In the \citeauthor{coq} proof assistant, for instance, inductive types, as well as functions and
theorems over inductive types, are closed to extension.
So to reuse mechanized metatheories,
the common practice is still to copy code and proofs and then modify them in each extension.
But having to maintain multiple copies is highly non-modular and
antithetical to good software engineering.
%
The programmer could also turn to design patterns for structuring developments
and tool support for cutting down on boilerplate
\cite{delaware2011,delaware2013,schwaab2013modular,keuchel2013generic,forsta2020}.
%\rD{The Coq à la Carte paper is arguably more than just a design pattern / encoding, given that it extends the surface syntax of the theorem prover. The same designation could ostensibly be applied to the plugin described in Sections 3-4. The difference, I suppose, is that the extended language in that work does not have a formal treatment.}
But these solutions tend to require heavy lifting from the programmer to make code
extensible and often lead to non-idiomatic programming styles.
